Covariate analysis of ROSMAP RNASeq

### 2016-01-26 code/Rmd/ROSMAP_RNASeq.Rmd
Added differential expression to the ROSMAP markdown

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

@@

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=10, fig.height=6}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom design)
Fitting a linear model with cogdx, Batch, RINcontinuous, and pmi 
```{r iterative.normalisation, cache=TRUE, results='asis', echo=FALSE}
postAdjustCovars <- c('cogdx','Batch',"RINcontinuous","pmi")

@@

```{r get.gene.symbols, cache=TRUE}
# Define biomart object
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ROSMAP RNASeq (Based on cogdx)", parentId = parentId)
CODE <- synStore(CODE)

>>>>

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(ggplot2)

@@

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=10, fig.height=6}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom design)
Fitting a linear model with cogdx, Batch, RINcontinuous, and pmi 
```{r iterative.normalisation, cache=TRUE, results='asis', echo=FALSE}
postAdjustCovars <- c('cogdx', 'Batch', 'RINcontinuous', 'pmi')

@@

```{r get.gene.symbols, cache=TRUE, eval=F}
# Define biomart object
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Fit contrast
contrast = makeContrasts(contrasts=c("cogdx1-cogdx2",
                                     "cogdx1-cogdx4",
                                     "cogdx2-cogdx4"), 
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = list()
p = list()
for (i in 1:3){
  DE[[i]] = topTable(FIT.CONTR, coef=i, number = dim(EXPR)[1])
  DE[[i]] = rownameToFirstColumn(DE[[i]], "ensembl_gene_id") 
  p[[i]] <- ggplot(DE[[i]], aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
  p[[i]] <- p[[i]] + ggtitle(colnames(contrast)[i])# + xlim(c(-1.5,1.5)) + ylim(c(0,6))
  p[[i]] <- p[[i]] + geom_hline(yintercept = -log10(0.05), color = 'red')
}
multiplot(plotlist=p, cols = 3)
names(DE) = colnames(contrast)

DE = DE %>% plyr::ldply(.id = 'Comparison')

tmp = DE %>%
  dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(Comparison) %>%
  group_by(Comparison) %>%
  dplyr::summarise(nDiffExp = length(unique(hgnc_symbol)), 
            DiffExpGenes = paste(unique(hgnc_symbol), collapse = ','))

kable(tmp)
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ROSMAP RNASeq (Based on cogdx)", parentId = parentId)
CODE <- synStore(CODE)
